---
title: Are You in the Room? A Framework for Sortition and Representation
author: Daniel Loza
date: today
toc: false
format: # Use only one of these
  presentation-revealjs+letterbox: default
---

```{r}
#| label: load-packages
#| include: false
library(knitr)
library(ggplot2)
library(dplyr)
library(readr)
library(tidyr)
library(tsibble)
library(feasts)
library(fable)
library(UpSetR)
library(here)
library(tidyverse)
library(devtools)
require(ggmosaic)
library(patchwork)
options(width = 70)
```

# Are you in the room?

## Sortition and representation



## The process

- Define a population and traits of interest
- Request expressions of interest from a random subset of the population
- Feed respondents into the sortition algorithm
- Use the generated panel for deliberation and decision-making

# But where do we assess the representativeness?

## The algorithms

LexiMin, or Lexicographic minimization

## What's new?

- Visualising to improve transparency
- Measuring performance and how extreme cases are handled
- Building a dashboard and framework to encourage evaluation of resulting panels against starting populations or prior stages

##

# Tables and graphs

## Slide 2

A nice table

```{r}
mtcars |>
  select(mpg:wt) |>
  head() |>
  kable(booktabs = TRUE)
```

## Slide 3

And a graph

```{r}
#| fig-height: 3.8
mtcars |>
  ggplot(aes(x=wt, y = mpg)) +
  geom_point()
```

## Slide 4

The sample standard deviation is given by
$$
  s = \sqrt{\frac{1}{n-1} \sum_{i=1}^n (x_i - \bar{x})^2}
$$

## Testing monospace font

```{r}
#| echo: true
mtcars |> as_tibble() |> head()
```


## Testing math fonts

* Calligraphic

  * $\mathcal{ABCDE}$

* Script

  * $\mathscr{ABCDE}$

* Fraktur

  * $\mathfrak{ABCDE}$

* Mathbb

  * $\mathbb{ABCDE}$
